### 2017-01-28 Radio/VLA_OH/VLA_2.tex
added captions, send to DR

\noindent \textbf{Science goals}
\vspace{.2em}\\
\indent {\bf Testing OHM as a tracer of merger}
The proposed \obs will establish the first OHM measurement in a {\it bona-fide} high-$z$ merger, thereby
demonstrating the feasibility of using it as a tracer of distant mergers and 
testing the long-envisaged hypothesis of using its number density to constrain 
the redshift evolution of merger rate. 
Such a detection will also allow us to better constrain the
\LOH-\LFIR relation out to higher $z$, % which may provide a handle on the pumping efficiency (\LOH/\LFIR) 
% that may be closer to other high-$z$ galaxies if their physical conditions are more similar than compared to low-z (\LOH/\LFIR)
and thus improve predictions on the required sensitivity needed for future searches of high-$z$ OHMs.
Since our target is a merger, 
a null detection would also indicate that the notion of 
using the OHM number density to constrain the merger history requires revision before designing future OHM surveys.

%
% OHM help us understand the physical conditions of high-z galaxies: different circumnuclear environment, optical depth? - likely already know about the cirucmnuclear env part from X-ray, but then could be use for other high-z sources w/o detailed characterization?
{\bf Physical conditions and OH pumping mechanisms in low- and high-$z$ galaxies}
% Classical model: foreground amplification of background radio continuum, expect quadratic relationship, if the population inversion is pumped by the FIR radiation field and the radio continuum (which itself is proportional to the FIR luminosity) is being amplified by an unsaturated maser cloud.
Null detections in previous high-$z$ OHM searches suggest a shallower \LOH-\LFIR dependence, which may be
reflecting the different physical conditions of the inner few hundred pc (where OHMs are generated) 
between low- and high-$z$ IR-luminous galaxies.
The lower \LOH/\LFIR indicate that the OH amplification is less efficient at the same FIR luminosity 
and that collisional pumping (e.g. due to the strong AGN, and high gas mass and density; \citealt{Field94a}) % volume density 
% while locally, not all OHM are AGN hosts, suggesting the FIR radiation from the SB is the dominant pumping source, but the non-detection based on local extrapolation and such maser model may suggest a different energy source/powering mechanism at higher-z, where quasars are more active, and may indicate that something about AGN/radio pumping
%
% Since a purely radiative pumping mechanism is possible 
% only at relatively low densities ($n_{\rm OH}$\ssim10$^{-4}$\,cm$^{-3}$), and collisional processes can theoretically predominate at high OH densities, and saturated amplification can become important \citep{Field94a}.
may play a dominant role in these gas-rich \highz systems.
Obtaining a definitive OHM measurement in RXJ1131 will provide a handle on the OH pumping efficiency (\LOH/\LFIR) out to higher $z$.
This together with the detailed properties derived for our target (e.g., \LFIR, \LIR, SFR, $L_{\rm radio}$, molecular and dust content)
will elucidate the role of collisional over radiative pumping
that may be prevalent in \highz galaxies given their physical properties, which are more similar to RXJ1131 than 
local starbursts. % more violent star formation in a more gas-rich turbulent ISM.
	% [NOT applicable] Since the optical depth increases with FIR radiation field, the simple picture of low-gain amplification
	% (i.e., unsaturated masing) found to be consistent with local OHMs...
	%
% % implying the simple picture of low-gain amplification (i.e., unsaturated masing, quadrature, low density)  \citep{Henkel90a,Randell95a} found to be consistent with local OHMs will likely break down; (i.e., not all OHMs are unsaturated). 
% While in the nearby universe (z\ssim0), OHMs appear to be unsaturated.

{\bf OHM as a tracer of the kinematics in the innermost region of distant galaxies}
The much broader linewidths found in $z$$>$0.2 OHMs compared to those at $z$\eq0 already indicate 
different dynamical properties in the innermost regions of their hosts \citep{Baan92a,Darling02a}.
Detecting the OHM in RXJ1131 will therefore demonstrate the feasibility of 
using it as a trace of the kinematics in the circumnuclear regions of \highz galaxies, which will improve 
our understanding of the internal dynamics of IR-luminous, starbursting galaxies at high $z$.

{\bf Radio continuum}
In addition, the proposed \obs will provide an additional constraint on the radio continuum emission at 1\,GHz 
towards RXJ1131 (\Fig{hst}e).

@@

\begin{figure}[ptbh]
\includegraphics[trim=10 305 270 0, clip, scale=0.825]{Fig/OHFIR} \hspace{-1.25em}
% \includegraphics[trim=0 400 200 0, clip, scale=0.85]{Fig/Darling02b_distribution}
%  \includegraphics[trim=0 0 0 0, clip, scale=0.625]{Fig/Willett11_distribution}\\ \vspace{1.5em}
% \includegraphics[trim=0 0 0 0, clip, scale=0.42]{Fig/Darling02_predictionf4}
\includegraphics[trim=0 -20 155 0, clip, scale=0.41]{Fig/willett12_fig1_mod.pdf}
\caption{{\bf \LOH-\LFIR correlation and existing OHMs measurements}
{\it (a):} Since the OH line strength increases with \LFIR,
the ubiquity of IR-luminous galaxies at high $z$ suggest that OHM can be used to trace distant mergers 
($L_{\rm IR}$$\gtrsim$10$^{12}$\Lsun; DG02).
{\it (b):} The highest-$z$ OHM known ($z$\eq0.265) was discovered over two decades ago. 
A remarkably low detection rate of $\sim$7\% was reported in subsequent surveys (black and green symbols), 
suggesting a potential redshift evolution in the \LOH/\LFIR relation
due to the different physical properties between low- and high-$z$ IR-luminous galaxies.
Assuming $L_{\rm OH}\propto L_{\rm FIR}^{1.2}$ (Malmquist-bias corrected; DG02), 
we expect an apparent OH luminosity of \LOH= 3.3\E{3}\,\Lsun in our target (orange marker). (Figure adopted from 
\citealt{Willett12a})
Here we propose to observe OH 1667 MHz line emission in the gravitationally-lensed starbursting 
galaxy merger RXJ1131 at $z$\eq0.65.
The proposed sensitivity will
probe deeper than existing OHM surveys, and to fainter intrinsic (i.e., lensing-corrected) \LOH beyond $z$$\gtrsim$0.2.
The proposed \obs will measure the \LOH/\LFIR 
and its redshift evolution, and thus improve model predictions for designing 
future surveys to search for high-$z$ OHMs with upcoming facilities. 
\label{fig:model}}
\end{figure}

\begin{figure}[ptbh]
% \hspace{-1.6em}
\centering
\includegraphics[trim=0 0 300 0, clip, scale=.75]{Fig/Fig2} 
\vspace{-0.25em}
\caption{
{\bf Our target RXJ1131, a gas-rich, starbursting merger at $z$\,$>$\,0.6 with a well-sampled spectral energy distribution (SED) spanning rest-frame UV to radio.} 
{\em (a):} Source-plane reconstruction of {\it HST} images shows recent star formation in our target and
the presence of a nearby companion (\citealt{Claeskens06a}).
{\em (b):} 
CO line emission overlaid on an {\it HST} image, which shows BLAH BLAH
that RXJ1131 is lensed into an Einstein ring of diameter $\sim$3.6".
{\em (c):}
Our CO \obs reveal large molecular gas reservoir of $M_{\rm gas}$\,$\gtrsim$\,10$^{10}$\Msun in RXJ1131 and 
confirmed the presence of the nearby companion, with a gas mass ratio of 7:1 \citep{Leung17a}. 
{\em (d):}
Our VLA radio continuum image at 5\,GHz in this target. 
The proposed \obs will provide an additional constraint on the radio continuum at 1\,GHz.
{\em (e):} Our best-fit model to the well-sampled SED 
indicates a strong apparent far-IR radiation of \LFIR$>$4\E{12}\,\Lsun, corresponding to
an intense star formation rate of SFR\ssim120 \Msun yr\pmOne (lensing-corrected).
The large molecular gas reservoir, on-going star formation, and the high \LFIR together
are favourable for generating OHM, and thus RXJ1131 is an
ideal candidate for high-$z$ OHM search.
The proposed \obs together with our existing data acquired for this target will allow us to 
compare its (host) properties against those of local OHM hosts to better understand 
the physical environment that triggers OHM out to a look-back time of 6 Gyr.
\label{fig:hst}}
\end{figure}

>>>>

\noindent \textbf{Science goals}
\vspace{.2em}\\
$\bullet$ {\bf Testing OHM as a tracer of merger}
The proposed \obs will establish the first OHM measurement in a {\it bona-fide} high-$z$ merger, thereby
demonstrating the feasibility of using it as a tracer of distant mergers and 
testing the long-envisaged hypothesis of using its number density to constrain 
the redshift evolution of merger rate. 
Such a detection will also allow us to better constrain the
\LOH-\LFIR relation out to higher $z$, % which may provide a handle on the pumping efficiency (\LOH/\LFIR) 
% that may be closer to other high-$z$ galaxies if their physical conditions are more similar than compared to low-z (\LOH/\LFIR)
and thus improve predictions on the required sensitivity needed for future searches of high-$z$ OHMs.
Since our target is a merger, 
a null detection would also indicate that the notion of 
using the OHM number density to constrain the merger history requires revision before designing future OHM surveys.

%
% OHM help us understand the physical conditions of high-z galaxies: different circumnuclear environment, optical depth? - likely already know about the cirucmnuclear env part from X-ray, but then could be use for other high-z sources w/o detailed characterization?
\noindent $\bullet$ {\bf Physical conditions and OH pumping mechanisms in low- and high-$z$ galaxies}
% Classical model: foreground amplification of background radio continuum, expect quadratic relationship, if the population inversion is pumped by the FIR radiation field and the radio continuum (which itself is proportional to the FIR luminosity) is being amplified by an unsaturated maser cloud.
Null detections in previous high-$z$ OHM searches suggest a shallower \LOH-\LFIR dependence, which may be
reflecting the different physical conditions of the inner few hundred pc (where OHMs are generated) 
between low- and high-$z$ IR-luminous galaxies.
The lower \LOH/\LFIR indicate that the OH amplification is less efficient at the same FIR luminosity 
and that collisional pumping (e.g. due to the strong AGN, and high gas mass and density; \citealt{Field94a}) % volume density 
% while locally, not all OHM are AGN hosts, suggesting the FIR radiation from the SB is the dominant pumping source, but the non-detection based on local extrapolation and such maser model may suggest a different energy source/powering mechanism at higher-z, where quasars are more active, and may indicate that something about AGN/radio pumping
%
% Since a purely radiative pumping mechanism is possible 
% only at relatively low densities ($n_{\rm OH}$\ssim10$^{-4}$\,cm$^{-3}$), and collisional processes can theoretically predominate at high OH densities, and saturated amplification can become important \citep{Field94a}.
may play a dominant role in these gas-rich \highz systems.
Obtaining a definitive OHM measurement in RXJ1131 will provide a handle on the OH pumping efficiency (\LOH/\LFIR) out to higher $z$.
This together with the detailed properties derived for our target (e.g., \LFIR, \LIR, SFR, $L_{\rm radio}$, molecular and dust content)
will elucidate the role of collisional over radiative pumping
that may be prevalent in \highz galaxies given their physical properties, which are more similar to RXJ1131 than 
local starbursts. % more violent star formation in a more gas-rich turbulent ISM.
	% [NOT applicable] Since the optical depth increases with FIR radiation field, the simple picture of low-gain amplification
	% (i.e., unsaturated masing) found to be consistent with local OHMs...
	%
% % implying the simple picture of low-gain amplification (i.e., unsaturated masing, quadrature, low density)  \citep{Henkel90a,Randell95a} found to be consistent with local OHMs will likely break down; (i.e., not all OHMs are unsaturated). 
% While in the nearby universe (z\ssim0), OHMs appear to be unsaturated.

\noindent $\bullet$ {\bf OHM as a tracer of the kinematics in the innermost region of distant galaxies}
The much broader linewidths found in $z$$>$0.2 OHMs compared to those at $z$\eq0 already indicate 
different dynamical properties in the innermost regions of their hosts \citep{Baan92a,Darling02a}.
Detecting the OHM in RXJ1131 will therefore demonstrate the feasibility of 
using it as a trace of the kinematics in the circumnuclear regions of \highz galaxies, which will improve 
our understanding of the internal dynamics of IR-luminous, starbursting galaxies at high $z$.

\noindent $\bullet$ {\bf Radio continuum}
In addition, the proposed \obs will provide an additional constraint on the radio continuum emission at 1\,GHz 
towards RXJ1131 (\Fig{hst}e).

@@

\begin{figure}[ptbh]
\includegraphics[trim=10 305 270 0, clip, scale=0.825]{Fig/OHFIR} \hspace{-1.25em}
% \includegraphics[trim=0 400 200 0, clip, scale=0.85]{Fig/Darling02b_distribution}
%  \includegraphics[trim=0 0 0 0, clip, scale=0.625]{Fig/Willett11_distribution}\\ \vspace{1.5em}
% \includegraphics[trim=0 0 0 0, clip, scale=0.42]{Fig/Darling02_predictionf4}
\includegraphics[trim=0 -20 155 0, clip, scale=0.41]{Fig/willett12_fig1_mod.pdf}
\caption{{\bf \LOH-\LFIR correlation and existing OHMs measurements}
{\it (a):} Since the OH line strength increases with \LFIR,
the ubiquity of IR-luminous galaxies at high $z$ suggest that OHM can be used to trace distant mergers 
($L_{\rm IR}$$\gtrsim$10$^{12}$\Lsun; DG02).
{\it (b):} The highest-$z$ OHM known ($z$\eq0.265) was discovered over two decades ago. 
A remarkably low detection rate of $\sim$7\% was reported in subsequent surveys (black and green symbols), 
suggesting a potential redshift evolution in the \LOH/\LFIR relation
due to the different physical properties between low- and high-$z$ IR-luminous galaxies.
Assuming $L_{\rm OH}\propto L_{\rm FIR}^{1.2}$ (Malmquist-bias corrected; DG02), 
we expect an apparent OH luminosity of \LOH= 3.3\E{3}\,\Lsun in our target (orange marker). (Figure adopted from 
\citealt{Willett12a})
Here we propose to observe OH 1667 MHz line emission in the gravitationally-lensed starbursting 
galaxy merger RXJ1131 at $z$\eq0.65.
The proposed sensitivity will
probe deeper than existing OHM surveys, and to fainter intrinsic (i.e., lensing-corrected) \LOH beyond $z$$\gtrsim$0.2.
Detecting the OHM in RXJ1131 will provide the first measurement beyond $z$$>$0.265,
which will improve predictions on the sensitivity needed for future high-$z$ OHM surveys with upcoming facilities.
\label{fig:model}}
\end{figure}

\begin{figure}[ptbh]
% \hspace{-1.6em}
\centering
\includegraphics[trim=0 0 300 0, clip, scale=.75]{Fig/Fig2} 
\vspace{-0.25em}
\caption{
{\bf Our target RXJ1131, a gas-rich, starbursting merger at $z$\,$>$\,0.6 with a well-sampled spectral energy distribution (SED) spanning rest-frame UV to radio.} 
{\em (a):} Source-plane reconstruction of {\it HST} images shows recent star formation in our target and
the presence of a nearby companion (\citealt{Claeskens06a}).
{\em (b):} 
The UV starlight in RXJ1131 is lensed into an Einstein ring (diameter $\sim$3.6") and the central AGN is 
quadruply-lensed in the {\it HST} image (grayscale). Both star formation and AGN are sustained 
by the extended CO gas reservoir \citep[contours; ][]{Leung17a}. 
{\em (c):}
Our CO \obs reveal a massive molecular gas reservoir of $M_{\rm gas}$\,$\gtrsim$\,10$^{10}$\Msun in RXJ1131 and 
confirmed the presence of the nearby companion, with a gas mass ratio of 7:1 \citep{Leung17a}. 
{\em (d):}
We have detected strong radio continuum at 5\,GHz towards this target. 
The proposed \obs will provide an additional constraint at 1\,GHz (orange marker in \Fig{hst}e).
{\em (e):} Our best-fit SED model indicates a strong apparent far-IR radiation of \LFIR$>$4\E{12}\,\Lsun, corresponding to
an intense star formation rate of SFR\ssim120 \Msun yr\pmOne (lensing-corrected).
The massive gas reservoir, on-going star formation with high \LFIR, and the large gas column density
are favourable for generating OHM, making RXJ1131 an
ideal candidate to search for OHM beyond the nearby universe.
Here we aim to detect the OHM in RXJ1131, and thus obtain a handle on the pumping efficiency
to better understand the triggering mechanism at a look-back time of 6 Gyr, which 
would reflect the different physical conditions of the innermost regions between low- and high-$z$ IR-luminous galaxies.
The proposed \obs is thus an important step towards understanding the evolution of the inner few hundred pc of
IR-luminous galaxies.
\label{fig:hst}}
\end{figure}
